A Comparison of Parallel Graph Processing Implementations

### 2017-05-15 papers/publication/paper-pollard.tex
minor

We aim to simplify the decision making process by providing an easy-to-use framework for consistently and fairly evaluating the performance of different algorithms. To illustrate the use of this framework, we include results from our experiments on both real world and synthetic datasets. We take inspiration from the Graph500 benchmark~\cite{Murphy:2010:Graph500}, which clearly specifies every step of the breadth first search (BFS) algorithm and how it should be timed. The Graph500 can be used to rank systems fairly on a single, well-defined benchmark. In this context, \emph{system} is defined as the hardware, operating system, middleware, and algorithmic implementation which accomplishes a certain task, in this case BFS.

@@

In addition, we analyze  power usage, energy consumption, and details such as the number of iterations for PageRank or the time to construct the graph data structures. Moreover, our approach requires little knowledge of the inner workings of each system; the data are collected by either parsing log files (for execution time) or hardware counter sampling of model-specific registers (for power). This allows arbitrary datasets to be included as long as they are in the same, simple format as those in the Stanford Network Analysis Project (SNAP) datasets~\cite{snapnets}.

@@

The difficulty in comparing iteration counts for PageRank underscores an important challenge for any comparison of graph processing systems. The assumptions under which the various platforms operate can have a dramatic effect on the program. For example, the GAP Benchmark Suite can be recompiled to store weights as integers or floating point values. This may affect performance in addition to runtime behavior in cases where weights like $0.2$ are cast to $0$. Similarly, how a graph is represented in the system (e.g., weighed or directed) may have performance and algorithmic implications but is not always readily apparent.

@@

These plots show generally poor scaling for this size problem, a challenge facing most parallel graph algorithms. 
\todo[inline]{If time permits, replace with a larger graph to show scaling.}
Additionally, $2^{20}$ vertices is a relatively ``small'' graph resulting for by today's standards and thus library designers focus on scalability for larger graphs. Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine. GCC version 4.8.5 supports OpenMP version 3.1 but the most recent release of OpenMP is 4.5.

\subsection{Real World Datasets}
\begin{figure}
	\centering
		% trim=left lower right upper
		\includegraphics[width=\linewidth, trim=0 18pt 6pt 0pt, clip]{graphics/compare-realworld.pdf}
	\caption{Real world experiments using \emph{easy-parallel-graph-\textasteriskcentered}. The leftmost plot does is missing Powergraph because Powergraph does not provide BFS.}
	\label{fig:epg-realworld}
\end{figure}

\begin{figure}
	\includegraphics[width=\linewidth, trim=0 18pt 6pt 0pt, clip]{graphics/graphalytics-GraphBIG.png}
	\caption{Real world and synthetic experiments using Graphalytics on GraphBIG using 32 threads. Graphalytics outputs one HTML page per software package.}
	\label{fig:graphalytics-html}
\end{figure}
%\begin{table}
%	\centering
%	\caption{This table shows averages of runtimes for the Dota-League dataset using our method with 32 threads. We average over 32 runs and since the data structure construction times do not differ significantly we average across all algorithms. Both PowerGraph and GraphBIG read in the files and build a graph data structure simultaneously so we cannot measure data structure construction time.}
%	\begin{tabular}{l|c|c|c|c}
%		\multicolumn{2}{l}{\large{\emph{easy-parallel-graph\mbox{-\textasteriskcentered}}}} & & &\\
%		System & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular}
%		& \begin{tabular}[x]{@{}c@{}}Run \\ BFS\end{tabular}
%		& \begin{tabular}[x]{@{}c@{}}Run \\ SSSP\end{tabular}
%		& \begin{tabular}[x]{@{}c@{}}Run \\ PageRank\end{tabular}
%		\\ \hline
%		GraphBIG   & N/A   & 0.01133        & {\bf 0.03791} & 19.5325  \\
%		PowerGraph & N/A   & N/A            & 0.3687        & 3.4406 \\
%		GAP        & 0.351 & 0.07449        & 0.0746        & {\bf 0.2465} \\
%		GraphMat   & 2.207 & {\bf 0.02273 } & 0.0397        & 1.6952 \\
%	\end{tabular}
%	\label{tab:epg-perf}
%\end{table}
%1         GAP      BFS Data structure build  0.25732000
%2    GraphMat      BFS Data structure build  2.10878359
%3         GAP PageRank Data structure build  0.32198688
%4    GraphMat PageRank Data structure build  2.15872441
%5         GAP     SSSP Data structure build  0.47479500
%6    GraphMat     SSSP Data structure build  2.35433178
%7         GAP      BFS                 Time  0.07449250
%8    GraphBIG      BFS                 Time  0.01132847
%9    GraphMat      BFS                 Time  0.02272806
%10        GAP PageRank                 Time  0.24646406
%11   GraphBIG PageRank                 Time 19.53251875
%12   GraphMat PageRank                 Time  1.69520531
%13 PowerGraph PageRank                 Time  3.44062500
%14        GAP     SSSP                 Time  0.07461594
%15   GraphBIG     SSSP                 Time  0.03791720
%16   GraphMat     SSSP                 Time  0.03936912
%17 PowerGraph     SSSP                 Time  0.36874997

@@

One overarching issue with these software packages is the speed at which they change. On one end of the spectrum, the code base of PowerGraph was made closed-source in 2015. Simple fixes such as increasing the maximum number of threads per process and updating dependency URLs are not merged and as a result there is much duplicated effort among more than 400 forks of PowerGraph. On the other end, GraphMat's update to version 2.0 is in general incompatible with version 1.0 and as such has not yet been adopted by Graphalytics nor our framework. One potential solution would be to add each framework to a package management system such as Spack \cite{Gamblin:2015:Spack} which supports packaging of high performance computing software.

@@

Overall, the GAP Benchmark Suite was the best-performing system across the given datasets though not the most scalable for the graphs used in this study (at most $2^{22}$ vertices). In addition, GAP is the most recent project and also the most robust in terms of the input graphs accepted. Thus, we recommend graph processing algorithm designers compare their BFS implementations to the Graph500 for BFS, GAP Benchmark Suite for smaller problem sizes or a low number of threads (at most 32), and GraphMat for larger problems with more threads. This is despite the fact that GraphBIG has the best scalability, because even at more than 16 threads the runtime for GraphBIG is still at least one order of magnitude slower than GAP or GraphMat for all sampled algorithms. Last, but not least, one must consider cost and portability: GraphMat requires the Intel compiler collection, which is not free or universally available for all platforms.

>>>>

We aim to simplify the decision making process by providing an easy-to-use framework for consistently and fairly evaluating the performance of different algorithms. To illustrate the use of this framework, we include results from our experiments on both real-world and synthetic datasets. We take inspiration from the Graph500 benchmark~\cite{Murphy:2010:Graph500}, which clearly specifies every step of the breadth first search (BFS) algorithm and how it should be timed. The Graph500 can be used to rank systems fairly on a single, well-defined benchmark. In this context, \emph{system} is defined as the hardware, operating system, middleware, and algorithmic implementation which accomplishes a certain task, in this case BFS.

@@

In addition, we analyze  power usage, energy consumption, and details such as the number of iterations for PageRank or the time to construct the graph data structures. Moreover, our approach requires little knowledge of the inner workings of each system; the data are collected by either parsing log files (for execution time) or hardware counter sampling of model-specific registers (for power). This allows arbitrary datasets to be included as long as they are in the same format as those in the Stanford Network Analysis Project (SNAP) datasets~\cite{snapnets}.

@@

The difficulty in comparing iteration counts for PageRank underscores an important challenge for any comparison of graph processing systems. The assumptions under which the various platforms operate can have a dramatic effect on the program. For example, the GAP Benchmark Suite can be recompiled to store weights as integers or floating-point values. This may affect performance in addition to runtime behavior in cases where weights like $0.2$ are cast to $0$. Similarly, how a graph is represented in the system (e.g., weighed or directed) may have performance and algorithmic implications but is not always readily apparent.

@@

These plots show generally poor scaling for this size problem, a challenge facing most parallel graph algorithms. 
\todo[inline]{If time permits, replace with a larger graph to show scaling.}
In addition, $2^{20}$ vertices is a relatively ``small'' graph resulting for by today's standards and thus library designers focus on scalability for larger graphs. Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine. GCC version 4.8.5 supports OpenMP version 3.1 but the most recent release of OpenMP is 4.5.

\subsection{Real-World Datasets}
\begin{figure}
	\centering
		% trim=left lower right upper
		\includegraphics[width=\linewidth, trim=0 18pt 6pt 0pt, clip]{graphics/compare-realworld.pdf}
	\caption{Real world experiments using \emph{easy-parallel-graph-\textasteriskcentered}. The leftmost plot does is missing Powergraph because Powergraph does not provide BFS.}
	\label{fig:epg-realworld}
\end{figure}

\begin{figure}
	\includegraphics[width=\linewidth, trim=0 18pt 6pt 0pt, clip]{graphics/graphalytics-GraphBIG.png}
	\caption{Real-world and synthetic experiments using Graphalytics on GraphBIG using 32 threads. Graphalytics outputs one HTML page per software package.}
	\label{fig:graphalytics-html}
\end{figure}
%\begin{table}
%	\centering
%	\caption{This table shows averages of runtimes for the Dota-League dataset using our method with 32 threads. We average over 32 runs and since the data structure construction times do not differ significantly we average across all algorithms. Both PowerGraph and GraphBIG read in the files and build a graph data structure simultaneously so we cannot measure data structure construction time.}
%	\begin{tabular}{l|c|c|c|c}
%		\multicolumn{2}{l}{\large{\emph{easy-parallel-graph\mbox{-\textasteriskcentered}}}} & & &\\
%		System & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular}
%		& \begin{tabular}[x]{@{}c@{}}Run \\ BFS\end{tabular}
%		& \begin{tabular}[x]{@{}c@{}}Run \\ SSSP\end{tabular}
%		& \begin{tabular}[x]{@{}c@{}}Run \\ PageRank\end{tabular}
%		\\ \hline
%		GraphBIG   & N/A   & 0.01133        & {\bf 0.03791} & 19.5325  \\
%		PowerGraph & N/A   & N/A            & 0.3687        & 3.4406 \\
%		GAP        & 0.351 & 0.07449        & 0.0746        & {\bf 0.2465} \\
%		GraphMat   & 2.207 & {\bf 0.02273 } & 0.0397        & 1.6952 \\
%	\end{tabular}
%	\label{tab:epg-perf}
%\end{table}
%1         GAP      BFS Data structure build  0.25732000
%2    GraphMat      BFS Data structure build  2.10878359
%3         GAP PageRank Data structure build  0.32198688
%4    GraphMat PageRank Data structure build  2.15872441
%5         GAP     SSSP Data structure build  0.47479500
%6    GraphMat     SSSP Data structure build  2.35433178
%7         GAP      BFS                 Time  0.07449250
%8    GraphBIG      BFS                 Time  0.01132847
%9    GraphMat      BFS                 Time  0.02272806
%10        GAP PageRank                 Time  0.24646406
%11   GraphBIG PageRank                 Time 19.53251875
%12   GraphMat PageRank                 Time  1.69520531
%13 PowerGraph PageRank                 Time  3.44062500
%14        GAP     SSSP                 Time  0.07461594
%15   GraphBIG     SSSP                 Time  0.03791720
%16   GraphMat     SSSP                 Time  0.03936912
%17 PowerGraph     SSSP                 Time  0.36874997

@@

One overarching issue with these software packages is the speed at which they change. On one end of the spectrum, the code base of PowerGraph was made closed-source in 2015. Simple fixes such as increasing the maximum number of threads per process and updating dependency URLs are not merged and as a result there is much duplicated effort among more than 400 forks of PowerGraph. On the other end, GraphMat's update to version 2.0 is in general incompatible with version 1.0 and as such has not yet been adopted by Graphalytics or our framework. One potential solution would be to add each framework to a package management system such as Spack \cite{Gamblin:2015:Spack} which supports packaging of high performance computing software.

@@

Overall, the GAP Benchmark Suite was the best-performing system across all chosen datasets, although not the most scalable for the graphs used in this study (at most $2^{22}$ vertices). In addition, GAP is the most recent project and also the most robust in terms of the input graphs accepted. Thus, we recommend graph processing algorithm designers compare their BFS implementations to the Graph500 for BFS, GAP Benchmark Suite for smaller problem sizes or a low number of threads (at most 32), and GraphMat for larger problems with more threads. This is despite the fact that GraphBIG has the best scalability, because even at more than 16 threads the runtime for GraphBIG is still at least one order of magnitude slower than GAP or GraphMat for all sampled algorithms. Last, but not least, one must consider cost and portability: GraphMat requires the Intel compiler collection, which is not free or universally available for all platforms.
